---
title: "GSS challange More Data"
author: "Nathaly Ruiz"
date: "Fall 2024"
output: html_document
---



```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom??
Data comes from the GSS of Chicago, 2016 survey 

$H_O$ There is no difference in attitude on capital punishment depending on race.             
$H_A$ There is a difference in attitude on capital punishment depending on race.

cappun~race         
is Categorical~Categorical          
so we will use the helper file Categorical~Categorical          


# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent")

```
The barchart shows that on average that more white people tend to be in favor of capital punishment, black people tend to not favor capital punishment, and other races tend to favor capital punishment. 

## Now with The Most data (fewer variables gives more data points)



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```
The numerical results shows that 65.10% of white people favor capital punishment while 34.90% oppose it.       
42.63% of black people favor capital punishment while 57.37% oppose it. 
55.20% other races favor capital punishment while 44.80% oppose it. 
This means that between the races white people favor capital punishment more than the other races, black people oppose capital punishment more than the other races, while other races average in attitude towards capital punishment between white and black people.


## Now with More data



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The inferential results shows the chi-squared test has a p-value of 0.00000000000000022, which is less than the acceptable value of 0.5 so we will reject the null hypothesis. The fishers exact test has a p-value of 0.00000000000000022, which is less than the acceptable rate of 0.5 so we will reject the null hypothesis.


## More data version


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
$H_O$ There is no difference in attitude on capital punishment depending on race.             
$H_A$ There is a difference in attitude on capital punishment depending on race.

Data comes from the GSS university of Chicago, 2016 survey.            
cappun~race         
is Categorical~Categorical          
so we will use the helper file Categorical~Categorical          
        
We will reject the null hypothesis because...          
The barchart shows that on average that more white people tend to be in favor of capital punishment, black people tend to not favor capital punishment, and other races tend to favor capital punishment.          
The numerical results shows that 65.10% of white people favor capital punishment while 34.90% oppose it.       
42.63% of black people favor capital punishment while 57.37% oppose it. 
55.20% other races favor capital punishment while 44.80% oppose it. 
This means that between the races white people favor capital punishment more than the other races, black people oppose capital punishment more than the other races, while other races average in attitude towards capital punishment between white and black people.             
The inferential results shows the chi-squared test has a p-value of 0.00000000000000022, which is less than the acceptable value of 0.5 so we will reject the null hypothesis. The fishers exact test has a p-value of 0.00000000000000022, which is less than the acceptable rate of 0.5 so we will reject the null hypothesis
